---
title: "exercises_week12"
author: "Lindsey Greenhill"
date: "4/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(stargazer)
library(gt)
library(skimr)
library(mice)
library(arm)
library(tidyverse)
set.seed(1121)
```

# Question 7.1

## Part a

```{r}

# write function for taking shots

shoot <- function(){
 shot_1 <- rbinom(n = 1, size = 1, prob = .6)
 bool <- FALSE
 count <- 1
 success <- if_else(shot_1 == 1, 1, 0)
 while(bool == FALSE){
   shot_2 <- rbinom(n = 1, size = 1, prob = .6)
   if((shot_2 == 0) & (shot_1 == 0)){
     count <- count + 1
     bool <- TRUE
   }
   else{
     success <- if_else(shot_2 == 1, success + 1, success)
     count <- count + 1
     shot_1 <- shot_2
   }
 }
 return(c(count,success))
}


```

## Part b

```{r}
# creating vector to store shot results

results <- tibble()

# simulating the shots 1000 times

for(i in 1:1000){
  r <- shoot()
  vec <- tibble(count = r[1],
                success = r[2])
  results <- results %>%
    rbind(vec)
}

# plotting the results

ggplot(results, aes(x = count)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Results of 1000 simulations",
       x = "Shots",
       y = "Count") +
  theme_classic()

# summary table of mean and sd

sum_tbl <- results %>%
  summarise(mean = mean(count),
            sd = sd(count))

sum_tbl %>%
  gt()

```

##  Part c

```{r}

# creating prop success col

results <- results %>%
  mutate(prop_success = success/count)

# scatterplot

ggplot(results, aes(x = count,  y = prop_success)) +
  geom_point() +
  theme_classic()
```

# Question 7.4

## Part a

```{r}

# reading in the data

beauty_data <- read.csv("http://www.stat.columbia.edu/~gelman/arm/examples/beauty/ProfEvaltnsBeautyPublic.csv", header = TRUE)

# cleaning the data

beauty_data <- beauty_data %>% 
  select(profevaluation, 
               female, courseevaluation, 
               btystdave, nonenglish, age, 
               profnumber)

# creating model

mod <- lm(courseevaluation ~ age + female + nonenglish + btystdave, beauty_data)

age <- c(50,60)
female <- c(1,0)
nonenglish <- c(0,0)
btstydave <- c(-1,-.5)

n.tilde <- length(age) # n.tilde is the number of predictions


X.tilde <- cbind(rep(1, n.tilde), age, female, nonenglish, btstydave)

head(X.tilde)

n.sims <- 1000
sim <- sim(mod, n.sims) # Make simulations


# creating sims

n.sims <- 1000
sim <- sim(mod, n.sims)

#sim.88 will give 1000 betas (similar to 1000 regression output) and 1000 sigmas.
head(sim@coef)
head(sim@sigma)

# Create an array to save the predicted results.

y.tilde <- array(NA, c(n.sims, n.tilde))

# Predict loop
for (s in 1:n.sims){
  y.tilde[s,] <- rnorm (n.tilde, 
                        X.tilde %*% sim@coef[s,], 
                        # multiply the new case matrix (X.tilde) by a matrix of simulated coefficients (sim.88@coef)
                        sim@sigma[s]
                        # extract the sigma for the predictions and use it as the standard deviation of rnorm()
                        )
}

dim(y.tilde)
head(y.tilde)


```

## Part b

In the code below, I created a histogram of the distribution of the difference
in course evaluation scores of A and B from the 1000 simulations created in part
a. I then calculated the probability teacher A's course evaluation score being
greater than teacher B's course evaluation score by calculating the proportion
of the difference values above 0.

** The probability that A has a higher score than B is 37.1% **

```{r echo=FALSE}

# creating data frame from sim results above. creating new column that is the
# difference between a and b where v1 is a and v2 is b. If diff is positive,
# then a had a better course evaluation than b.

new_df <- y.tilde %>% 
  as.tibble() %>%
  mutate(diff = V1 - V2)

hist <- new_df %>%
  ggplot(aes(x = diff)) +
  geom_histogram(binwidth = .1) +
  labs(title = "Difference in Course Evaluation Scores between A and B",
       subtitle = "1000 simulations",
       x = "Differencce in course evaluations") +
  theme_classic()

# calculating the probability that diff is greater than 0 by calculating the
# proportion of values above 0 in the distribution

prop <- new_df %>%
  mutate(greater = if_else(diff > 0, 1, 0)) %>%
  count(greater) %>%
  pivot_wider(id_cols = everything(),
              names_from = "greater",
              values_from = "n",
              names_prefix = "count") %>%
  summarise(prop = count1/(count0 + count1))

hist
```



