---
title: "exercises_week10"
author: "Lindsey Greenhill"
date: "4/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(contrasts = rep("contr.treatment",2))
save.image()
library(MASS)
library(stargazer)
library(gt)
library(tidyverse)
set.seed(123)
```

# Question 2

```{r include = FALSE}

# reading in data

nes <- foreign::read.dta("http://www.stat.columbia.edu/~gelman/arm/examples/nes/nes5200_processed_voters_realideo.dta") %>% 
  filter(year == 2000) %>%
  select(partyid3,age, gender, race, educ1, income, ideo7)

# recoding and cleaning data. Taken from a pset a while back. added the part of
# filtering out the apolitical party from the partyid3. Realized this is a
# mistake from my previous coding. Also, I had to keep it as numbers in order
# for the plot in part a to work.

nes_clean <- nes %>%
  drop_na() %>%
  filter(partyid3 != "9. apolitical (1966 only: and dk)") %>%
  mutate(
    gender = as.numeric(str_sub(gender, 1, 1)) - 1,
    educ1 = as.double(str_sub(educ1, 1, 1)),
    partyid3 = as_factor(str_sub(partyid3, 1, 1)),
    income = as.double(str_sub(income, 1, 1)),
    ideo7 = as.double(str_sub(ideo7, 1, 1)),
    race = as_factor(str_sub(race, 1, 1))
  ) %>%
  #mutate(partyid3 = if_else(partyid3 == 3, "R",
   #                         if_else(partyid3 == 2, "I", "D"))) %>%
  mutate(race = case_when(race == 1 ~ "white",
                          race == 2 ~ "black",
                          race == 3 ~ "asian",
                          race == 4 ~ "native_american",
                          race == 5 ~ "hispanic"))

# setting the levels to be Democrat, Independent, Republican

nes_clean$partyid3 <- factor(nes_clean$partyid3, levels = c(1,2,3))

nes_clean_simple <- nes_clean %>%
 select(-race, -gender)
```


## Part a

In the code below, I created both an ordered multinomial logit model (mod_1) and
probit model (mod_2). See the stargazer table below for the estimates for both
these models. For the dependent variable, partyid3, 1 = Independent, 2 =
Democrat, and 3 = Republican. The reference category is Independent.

```{r results = "asis"}

# creating multinomial logit (ordered). Chose to include race so the graphical
# plot below looks more normal

mod_1 <- polr(factor(partyid3) ~ age + educ1 + income + 
                ideo7 + race,
              data = nes_clean)

# creating probit model

mod_2 <- polr(factor(partyid3) ~ age + educ1 + income + ideo7 + race,
              data = nes_clean,
              method = "probit")

# summary of both models

# summary(mod_1)
# summary(mod_2)

stargazer(mod_1, mod_2, type = "latex")

```

In the code below, I create a graphical representation of the multinomial logit
model. The plot is based on figure 6.4 from the text book, where the curves show
the expected responses for party identification based off of ideology. The data
points do not seem to follow the curve very well, suggesting that the model
might not be the best fit.

```{r}

# graphs of both models. Make these like the ones in 6.4 of the textbook. Had to
# create the inverse logit function. Maybe I could have gotten this from the arm
# lib? Not sure

invlogit <- function(Xb) {
  1/(1+exp(-Xb))
}

# taken pretty much directly from the book

expected <- function (x, c1.5, c2.5, sigma){ 
  p1.5 <- invlogit ((x-c1.5)/sigma) 
  p2.5 <- invlogit ((x-c2.5)/sigma)  
  return ((1*(1-p1.5) + 2*(p1.5-p2.5) + 3*p2.5))
} 

# graph of the mod_1. We don't need to do the probit model for this part. It's a
# bit weird that the cut points are outside of the range of the data.

{
plot (nes_clean$ideo7, nes_clean$partyid3, xlim=c(0,12), ylim=c(1,3), 
      xlab="Ideology", ylab="Party")
lines (rep (mod_1$zeta[1]/mod_1$coefficients[4], 2), c(1,2)) 
lines (rep (mod_1$zeta[2]/mod_1$coefficients[4], 2), c(2,3))  
curve (expected (
  x, mod_1$zeta[1]/mod_1$coefficients[4],
  mod_1$zeta[2]/mod_1$coefficients[4],
  1/mod_1$coefficients[4]),
  add=TRUE) 
}

```


## Part b

Explain the results of the fitted model

I thought about the results of the fitted model by looking at the statistical
significance of each estimate and the value of that estimate relative to the
cutoff points of the ordered model

* Cut points: Democrat:Independent 4.8669, Independent:Republican 5.278

* Statistical significance: the educ1, income, and ideo7 coefficients are
statistically significant at the alpha = .05 level. The age and race variables
are not statistically significant.

* educ1 coefficient = .343: the education coefficient (.343) is less than the
difference between the cut points (~.41). This means that an increase in
education by itself will not actually move a voter from one political party to
another.

* income coefficient = .196: the income bracket coefficient (.196) is elss than
the difference between the cutoff points, similar to the education coefficient.
Thus, a one increase in income categories doesn't move a voter on average from
one party to another.

* ideo7 coefficient = .724: the ideology coefficient (.724) is greater than the
difference between the cut points, meaning that an increase on the ideology
scale has a relatively larger effect than the education or income scales and can
bring a voter from one party to another party on average.

# Question 3

### Data

I am using the rodents data from the week 7 pset. 

```{r include = FALSE}

# reading in the data

rodents <-
  read.delim("http://www.stat.columbia.edu/~gelman/arm/examples/rodents/rodents.dat",
             sep = " ")

# cleaning the data. Using code from old pset. 
# recoding the values so they will be easier to interpret.

rodents$race <-
  dplyr::recode(
    rodents$race,
    "1" = "White",
    "2" = "Black",
    "3" = "Puerto Rican",
    "4" = "Other Hispanic",
    "5" = "Asian/Pacific Islander",
    "6" = "Amer-Indian/Native Alaskan",
    "7" = "Two or More Races"
  )

# Not sure what categories would be appropriate to combine so I'm not going to
# combine any.

rodents <- rodents %>%
  mutate(race = as_factor(race))

# making white the first level in the factor variable

rodents$race <- relevel(rodents$race, "White")
```


## Part a

In the code below I create a logit (mod_3) and probit (mod_4) model. 

```{r results="asis"}

# logit model

mod_3 <- glm(rodent2 ~ race + dilap + regext , data = rodents,
             family = binomial(link = "logit"))

# probit model

mod_4 <- glm(rodent2 ~ race + dilap + regext , data = rodents,
             family = binomial(link = "probit"))

# displaying the two models

stargazer(mod_3, mod_4, type = "latex")
```


The logit and the probit estimates should be more or less the same after scaling
by 1.6. We can check this by multiplying the probit coefficients by 1.6. After
making this adjustments, we see that the estimates roughly match up with
eachother (see table below).

```{r}

# scaling the coefficients by 1.6

new_coefs <- mod_4$coefficients * 1.6

# making table of new coefficients. Using gt for better presentation

coefs_tb <- tibble(mod_3_est = mod_3$coefficients, 
       mod_4_est = mod_4$coefficients,
       mod_4_est_scaled_1.6 = new_coefs) %>%
  gt() %>%
  fmt_number(columns = everything(),
             decimals = 2)

# displaying table

coefs_tb
```

