---
title: "exercises_week13"
author: "Lindsey Greenhill"
date: "4/28/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(arm)
library(stargazer)
library(tidyverse)
set.seed(18)
```

# Question 11.4

```{r include=FALSE}

# reading in the data

data <- read.csv("http://www.stat.columbia.edu/~gelman/arm/examples/cd4/allvar.csv", 
                 header=TRUE)

# cleaning the data. making a root percentage

data <- data %>%
  drop_na() %>%
  mutate(sqrtpct = sqrt(CD4PCT),
         time = visage - baseage)

```

## Part a

In the code below, I graph the square root CD4 percentage against time for all
patients and for 16 randomly sampled patients. Both graphs show that there
doesn't appear to be a consistent pattern amongst the patients. Some patients
seems to have a negative correlation between CD4 levels and time while other
appear to have no correlation or positive correlation.

```{r echo=FALSE}

# not sure if we should just graph all of the data points or separate them by
# child?

graph_11_a_1 <- data %>%
  ggplot(aes(x = time, y = sqrtpct)) +
  geom_point(alpha = .5) +
  labs(title = "Sqrt CDC level vs Time",
       x = "Time",
       y = "Sqrt CDC") +
  theme_classic()

# should facet wrap


# take random sample

sample <- sample(unique(data$newpid), 16, replace = FALSE)

# filtering to random sample

data_11 <- data %>%
  filter(newpid %in% sample)

# graphing the individuals separately

graph_11_a_2 <- data_11 %>%
  ggplot(aes(x = time, y = sqrtpct)) +
  facet_wrap(~newpid) +
  geom_point() +
  geom_line() +
  labs(title = "Sqrt CD4 level vs time",
       subtitle = "Random Sample of 16 patients",
       x = "Time",
       y = "Sqrt CD4 level") +
  theme_classic()

graph_11_a_1
graph_11_a_2
```


## Part b

In the code below, I show the linear fit of sqrt CD4 level vs time for the 16
randomly sampled individuals from part a. As noted in part a, there does not
seem to be a consistent relationship between CD4 levels and time amongst these 16
patients.

```{r echo=FALSE, message=FALSE, warning=FALSE}

graph_11_b_1 <- data_11 %>%
  ggplot(aes(x = time, y = sqrtpct)) +
  facet_wrap(~newpid) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Sqrt CD4 level vs time Linear Regression",
       subtitle = "Random Sample of 16 patients",
       x = "Time",
       y = "Sqrt CD4 level") +
  theme_classic()
graph_11_b_1
```

## Part c

In the code below, I first calculate the slopes and intercepts of CD4 level
for each patient in the data. In the first step, I group the data by newpid
to generate intercepts and slopes specific to each indiviudal patient. In 
the next step, I regressed the intercept and slopes from the individual models
on baseage and treatment to get a 


```{r message=FALSE}

# creating regression coefficients grouped by newpid

c_mod <- data %>%
  group_by(newpid, baseage, treatmnt) %>%
  summarise(intercept = coef(lm(sqrtpct ~ time))[1],
         slope = coef(lm(sqrtpct ~ time))[2])

# creating models to explain the differences in between the children

c_mod_1 <- lm(intercept ~ baseage + treatmnt, data = c_mod)
c_mod_2 <- lm(slope ~ baseage + treatmnt, data = c_mod)

summary(c_mod_1)
summary(c_mod_2)

```

The first summary shows the regression output for the intercept model. The
coefficient for baseage means that the model predicts that on average for every
1 year increase in baseage there is a .12 point decrease in the intercept for
in between children holding all else constatn. The model predicts that that
treated children will on average have an intercept .25 points higher compared to
non treated children, holding all else constant.

The second summary shows the regression output for the slope model. The
coefficient for baseage means that the model predicts that on average for every
1 year increase in baseage a child's slope will be .02 points lower, holding all
else constant. The model predicts that that treated children will on average
have a slope .06 points lower compared to non treated children, holding all else
constant.

# Question 12.2

## Part a

In the code below, I create a model that predicts CD4 levels as a function of
time, varying intercepts across children.

```{r}
mod_12_a <- lmer(sqrtpct ~ time + (1 | newpid),
                 data = data)

# can't use stargzer with this model

summary(mod_12_a)
```


* Coefficient of time: -.38. For every year following the initial visit, the
patient's square root CD4 level is predicted to decrease by .38 on average. This
coefficient is statistically significant.

* Additionally, it seems like grouping the model by patient improves the model
because the sd of its random effects is larger than that of the residual.


## Part b

In the code below, I extend the model in part a to include more predictors,
specifically treatment and age at baseline.

```{r}

mod_12_b <- lmer(sqrtpct ~ time + treatmnt + baseage + (1 | newpid),
                   data = data)

summary(mod_12_b)

```

### Coefficient interpretations

* Coefficient of time: -.38. For every year following the initial visit, the
patient's square root CD4 level is predicted to decrease by .38 on average,
holding all else constant. This coefficient is statistically significant.

* Coefficient of base age: -.118. for every increase in the base age of a
patient, the model predicts that the square root CD4 level will decrease by .11
on average, holding all else constant. This coefficient is statistically
significant.

* Coefficient of treatment: .294. The model predicts that the patients who
received treatment have on average .294 higher square root CD4 levels than those
who did not receive the  treatment, holding all else constant. The coefficient
is not statistically significant.


## Part c and d

In the code below, I look at the y hats for the models made in part a and part b
above. I then plot these y hats against the only shared predictor in the models,
time since baseline visit. The predictions from model a and the predictions from
model b appear to be more or less the same. I confirmed this by looking at the
distribution of the differences of predictions between b and a in the histogram
below. The average difference in predictions is 0.

Additionally, we can compare the results from part a and b by looking at the
different sigma and alpha values. For the first model, sigma = 1.39 and alpha =
.76, meaning that the variation between groups is more than the variation within
groups. For the second model, sigma = 1.36, which is less than sigma in the
first model. This decrease suggests that adding the two predictors of baseage
and treatment reduces variation between groups.

```{r echo=FALSE}

# predictions for model a

pred_a <- as_tibble(predict(mod_12_a, newdata = data)) %>%
  rename(A = value)

# predictions for model b

pred_b <- as_tibble(predict(mod_12_b, newdata = data)) %>%
  rename(B = value)

# creating an x column

time <- data$time

# combining the data

comb <- pred_a %>% cbind(pred_b) %>%
  cbind(time)

# transforming the data into long format for plotting

comb_long <- comb %>%
  pivot_longer(cols = c("A", "B"),
               names_to = "model",
               values_to = "yhat")

# creating graph

graph_c <- comb_long %>%
  ggplot(aes(x = time, y = yhat, col = model)) +
  geom_point(alpha = .3) +
  labs(title = "Model Predictions: CD4 vs time",
       subtitle = "Models a and b have similar predictions",
       x = "Time",
       y = "y hat") + 
  theme_classic()

# difference in predictions

comparison <- comb %>% mutate(difference = B - A)

# histogram of differences

plot_c_2 <- comparison %>%
  ggplot(aes(x = difference)) +
  geom_histogram(binwidth = .01) +
  geom_vline(xintercept = mean(comparison$difference),
             color = "red") +
  labs(title = "Distribution of Prediction Differences",
       subtitle = "The avg difference between B and A is 0",
       x = "B yhat - A yhat") +
  theme_classic()

graph_c
plot_c_2
```











